The Talking Heads Experiment

### 2015-04-10 chap10/chap10.tex
fixes
quotation marks for chapters 8 to 10

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.50\textwidth]{chap10/figs/look-ball}
\includegraphics[width=.45\textwidth]{chap10/figs/views-ball}}
\caption{\label{f:look-ball} 
Left: Fr\'{e}d\'{e}ric Kaplan interacts with the AIBO robot in an experiment for the social learning of lexicons.
Right: View of the ball through the eye of the robot. As the same object is now seen from many different viewpoints it 
becomes much more challenging to consistently recognize it.}
\end{figure}
The first AIBO experiments did not involve a population of agents but rather a single agent that was interacting directly with a human
experimenter. The agent had to learn the lexicon of the human through social 
interaction within the context of situated language games. The 2001
paper `AIBO's first words' by Steels and Kaplan described experiments in which an AIBO robot 
learned words about objects in the environment, such as `ball' or `smiley', as well as words for actions, such as 
`sit down' or `stand up'.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.70\textwidth]{chap10/figs/fsm}}
\caption{\label{fig:fsm} 
The behaviour on the AIBO, including the language game scripts, are programmed using probabilistic finite state 
machines that bring the robot from one internal state to another one depending on internal or external conditions.}
\end{figure}

@@

The Perspective Reversal Experiment applied this methodology to explain a remarkable universal feature 
of human languages, namely, 
that speakers may use a different perspective on the scene than their own when 
conceptualizing what to say, and that they may explicitly mark perspective switching, using lexical or grammatical means. Suppose
a speaker is facing a hearer, then it indeed makes a big difference whether she says 
\emph{the door to your left}  or \emph{the door to my left}. If the speaker says
\emph{the door to my left} she expects the hearer to perform an egocentric
perspective transformation and see the situation from her own point of view. Perspective switching is here signalled by 
explicitly expressing which perspective is used: {\itshape your} left versus {\itshape my} left. 
The German prepositions \emph{herein} and \emph{hinein}, both meaning `into' are another example of perspective marking. They
distinguish whether the direction of movement is 
towards the speaker \emph{herein} or away from the speaker \emph{hinein}, analogous to English \emph{come} (to where I am) 
versus \emph{go} (to another location). Although there are obvious differences in how languages express
perspective, there can be no doubt that perspective marking is pervasive across 
languages and language families, and that speakers make abundant use of it.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap10/figs/tokyo-persp}}
\caption{\label{fig:tokyo-persp} 
Luc Steels demonstrates live the Perspective Reversal Experiment in Tokyo, April 2005. AIBO robots track 
the movement of a ball and 
the robot acting as speaker describes to the hearer an observed movement, possibly taking into account the 
perspective of the hearer. For example, the speaker might say `the ball rolls from my left to your right'.}
\end{figure}

The Perspective Reversal Experiment worked as follows. A population of AIBO robots roam around
freely in an unconstrained in-door environment. As soon as one sees the
ball, it comes to a stop and searches for another robot nearby, which also
looks for the ball and stops when it sees it. Then the human
experimenter pushes the ball with a stick so that it rolls a short
distance (see \figref{fig:tokyo-persp}). The two robots next play a description game. One robot (acting as the `speaker') 
describes the ball-moving event to the other robot (the `hearer'), going through the same semiotic cycle as the Talking 
Heads agents. The game is a success if the meaning conveyed by the speaker is compatible with the hearer's 
own perception of the scene.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.90\textwidth]{chap10/figs/persp-rev-perception}}
\caption{\label{fig:persp-rev-perception} 
Chain from visual image (left) to an analog world model (middle) and a discrete event description
that is the topic of the language game (right). 
The top row shows this chain for robot A and the bottom row the chain for robot B, both looking at the same ball.}
\end{figure}

Within this experimental setting three different strategies were tried in a long-term experiment where
a new strategy was injected at certain stages and performance monitored: 
\begin{itemize}
\item In stage I, agents use 
the Discrimination Game strategy and the Naming Game strategy discussed in chapters 4 and 5. A 
lexicon is emerging and there are some successful games (those for which the speaker and the hearer see the situation from 
the same perspective). But the lexicon keeps expanding and success is very limited as games where robots have a different 
perspective on the rolling ball all fail. 
\item In stage II, agents have an extended 
strategy where agents can geometrically transform a model of the scene, based on their own perspective, into a model of the scene 
as seen from the perspective of the hearer. They can then conceptualise the scene from that perspective. 
Communicative success significantly increases and the 
lexicon starts to decrease. However there is ambiguity because the hearer cannot know whether
the speaker has conceptualized the 
scene from his own viewpoint or performed a perspective reversal, leading to a large number of failed communications 
(more than 50 \%). 
\item In stage III, a more elaborate stratergy is used. The speaker now expresses whether he has performed perspective 
reversal (as in `{\itshape to your} left' vs. simply `left'). Only with this strategy are agents able to reach high 
communicative success with a stable inventory. 
\end{itemize}

\begin{figure}[t]
\centerline{
\includegraphics[width=0.75\textwidth]{chap10/figs/stages}}
\caption{\label{fig:stages} 
Three strategies are compared. In Stage I (left) agents categorise the event from their 
own perspective. In Stage II (middle) they conceptualise the scene from the viewpoint of the other agent as well. In Stage III (right) 
they express that they performed a perspective reversal. Only the third strategy leads to successful performance.} 
\end{figure}

@@

The earliest attempts to investigate the emergence of syntax date back to 1997, in other words even before 
the first Talking Heads experiment started. The first report 
appeared as a paper in the Artificial Intelligence journal in 1998 \citep{Steels:1998}. 
It was described by the journal editors (Dan Bobrow and 
Mike Brady) as `obviously in an early stage, but very provocative and suggestive.' The paper was based on my keynote 
lecture at IJCAI 1997 in Nagoya and used a precursor of the pan-tilt cameras built by Tony Belpaeme. 
(see \figref{f:plate7}).

@@

`The grammar is seen as a natural continuation of the lexicon, in the sense that it consists also of associations between forms and meanings. Use and success are monitored for each association so that the same type of self-organized coherence arises in the group, as seen in the lexicon. The form is now a more complex structure, defined as a syntactic schema. The meaning is a semantic schema. The schemas
circumscribe a feature-set in terms of a set of slots, restrictions	
on the fillers of each slot, and constraints	on the combination	
of the fillers to form the total covered by the schema.

@@

The categories restricting slot-fillers are either themselves defined in terms of schemas
(for example, syn-cat-77 could be the restriction on a slot-filler in another schema), or they
are defined as rules that are applied in a forward-chaining fashion during matching. Two
examples of rules related to the above schema are:
\begin{verbatim}
  rule 101: ([WORD (W U)]) => ([MEMBER syn-cat-761])
  rule 99: ([WORD (W O)]) => ([MEMBER Syn-Cat-751])
\end{verbatim}
Semantic schemas describe the language-specific semantic structures underlying the
meanings of complete word groups. The closest linguistic correspondent to a semantic
schema is the notion of a case-frame. The constraints indicate how the total meaning
is constructed/decomposed into the meaning of the parts. During interpretation such
constraints therefore perform the same role as Montague style semantic interpretation
functions. The slots correspond to cases such as agent, patient, time, distance, or arguments
of semantic functions. The categories used to constrain what can fill a slot correspond
in linguistic terms to selection restrictions like animate, human, edible, future, etc. The
schema has also an associated category for the whole so that hierarchical combination is
possible. An example of a semantic schema is:
\begin{verbatim}
Schema-542
  SLOTS: (sem-slot-51 sem-slot-50)
  DESCRIPTION-SET: 
       ([sem-slot-50 sem-cat-751] [sem-slot-51 sem-cat-761])
  CONSTRAINTS: 
       ((CONJUNCTION (>> sem-slot-50) (>> sem-slot-51)))
  CATEGORY: sem-cat-77
  USE: 10
  SUCCESS: 3
\end{verbatim}
Inference rules such as the following define the selection restrictions:
\begin{verbatim}
  rule 102: ([VISIB v-4111) => ([MEMBER sem-cat-761)
  rule 100: ([VER v-4311) => ([MEMBER sem-cat-751)
\end{verbatim}
Each association in the grammar couples a syntactic schema with a semantic schema.
The association can be used in two directions. If a syntactic schema is recognized, i.e., all its 
constraints are satisfied by the input utterance, the semantic schema is
used to reconstruct its meaning. If a semantic schema is recognized, the syntactic schema
is used to reconstruct the form. The association contains a mapping of the slots in order to
enable this reconstruction. For example, the association combining the above two schemas is: 
\begin{verbatim}
Association-271
  MEANING: Schema-542
  FORM : Schema-541
  MAPPING : ((syn-slot-51 sem-slot-51)
             (syn-slot-50 sem-slot-50))
  USE: 10
  SUCCESS: 3"
\end{verbatim}\citep{Steels:1998}

\begin{figure}[t]
\centerline{
\includegraphics[width=0.75\textwidth]{chap10/figs/associations}}
\caption{\label{fig:associations} 
Top: Two lexical associations have become active (association-259 and association-234) to cover the distinctive 
feature-set ([VISIB v-411] [VER v-431]) resulting from discrimination and perception. Bottom: 
The group of words and the meanings match with schemas forming part of a grammatical association that combines the words. }
\end{figure}
An application of these schemas in producing the utterance `(W O) (W U)' is represented in \figref{fig:associations} and shown 
here: 
\begin{verbatim}
127 ++> Speaker: head-40 Topic: 2 Context: 6 5 4 3 1 0
Categorial Perception:
  ([VISIB v-4111] [VER v-431])
  ([VER v-4311] [AREA v-4061])
Lexicon lookup: (Association-259 Association-234)
Syntactic structure:
  (syn-cat-77
    (syn-slot-50 (syn-cat-75 |(W O)|))
    (syn-slot-51 (syr-cat-76 |(W U)|)))
Semantic structure:
  (sem-cat-78
    (sem-slot-50 (sem-cat-75 (VER v-431)))
    (sem-slot-51 (sem-cat-76 (VISIB v-411))))
Meaning:
  ([VISIB v-4111] [VER v-4311)
Expression: ((W 0) (W U))
\end{verbatim}

@@

The paper also briefly described the mechanisms by which such a grammar emerged in the experiment in terms of a 
cognitive memory system: `The cognitive memory system acts in a first instant purely as a device that records 
a particular way in which language has been produced so that it can later be re-produced in the same way.
Also the hearer can perform this recording operation. He is presented with a specific set of word forms from which he can abstract a syntactic schema. He derives meaning from the definition of the words in the lexicon and the distinctive feature-set coming out of perception. From this, a semantic schema can be extracted.
More interesting grammars emerge when additional operations to restructure the set of grammatical associations are used. For example, in the case of a partial match, substructures may be recategorised so that they nevertheless fit, two partially overlapping schemes with common fillers may lead to a new schema that integrates both, etc.'

@@

Based on these components, I implemented an initial experiment with only two agents.\footnote{These experiments were first reported at the Harvard Evolution of Language conference in 2002 with a paper by Luc Steels
entitled `Computer Simulations of the origins of Case grammar'. The paper was later deemed 
`too incomprehensible for the Evolution of Language community' to be acceptable for the conference 
proceedings edited by Maggie Tallerman.}
In 2005 Remi van Trijp picked up this line of research again and worked out the experiment 
further with great skill and sophistication. He extended the 
2-agent simulation to a multi-agent simulation and could finally report a complete running system in his 2008 Ph.D 
thesis \citep{VanTrijp:2014}. Here I give only a brief account of the first attempts towards case grammar
and how this pushed the state of the art in language game research.

But first we need a clearer idea on what case grammar is.
In the sentence `John gave Mary a book', John plays the role of agent, the giver 
of the book, Mary that of recipient or 
beneficiary and the book is the `patient' or object of the action. Languages differ in terms of the strategies that they 
use to express these participant roles. Some of them, like German, have an elaborate system of morphological 
markers attached to constituents of the noun phrase. These markers signal that the 
phrase is nominative, accusative, dative, etc., and these cases then map in complex ways 
onto semantic roles such as agent and patient.
Other languages, like English, use a strategy based on word order and prepositions. For example, 
`Mary gave John a book' contains the same participants as `John gave Mary a book' but now Mary is the 
agent and John the recipient. Still other languages, like Japanese, use a strategy based on particles following 
the noun, as in the Japanese sentence `Tanaka-san wa Tokyo de o-to-san ni atta' 
meaning Tanaka met his father in Tokyo. Tanaka-san is the name of a person, oto-san means `father' and atta 
means `met'. The markers are `wa' for the topic, `de' for location and `ni' for another sense of location which can 
also function as possessive. The case grammar experiments focused on modeling the latter strategy.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.60\textwidth]{chap10/figs/peract-setup}}
\caption{\label{fig:peract-setup} 
The installation for the case grammar experiments used a more sophisticated vision system and objects that 
could be moved around. We see the two Talking Heads cameras in 
the background. They are oriented towards objects on the table or situations in the 
environment. The monitors show the camera images and the states of image processing, just as for the original 
Talking Heads experiment. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.40\textwidth]{chap10/figs/adam-eve}}
\centerline{\includegraphics[width=.70\textwidth]{chap10/figs/jack-give}}
\centerline{\includegraphics[width=.80\textwidth]{chap10/figs/push-jill}}
\caption{\label{fig:adam-eve} 
Top: Computer animation of the speaking and hearing agent used in public demonstrations 
of the case experiment. 
Middle: Example of a puppet scene, where Jill gives a block to Jack. It shows the source image (left) and a (partial) 
result of image processing (right). Bottom: Another example of a puppet scene where Jill pushes the block.}
\end{figure}
\\
{\bfshape  Scaling up the vision system} 
The vision system of the Talking Heads only dealt with static scenes. This had to be scaled up to
deal with dynamical scenes and event recognition. It was the focal work of 
Jean-Christophe Baillie at the Sony Computer Science Laboratory in Paris, who built a vision system 
called PERACT \is{PERACT vision system} inspired by an event-recognition developed by Jeffrey 
Siskind.\footnote{The vision system is described in the Ph.D thesis of J-C Baillie. A summary is found here: 
\cite{Steels:2003}. Inspiration for the event recognition system came 
primarily from the paper by \cite{Siskind:2000}.}
The vision system was decomposed into three subsystems with information flowing both in a bottom-up
(from image to world model) and top-down way (from world model and predictions to image): the subsystems achieved 
object identification, event identification, and derivation of qualitative descriptors.

{\bfshape  Object Identification} The first subsystem attempts to detect and track visual units at different hierarchical levels. 
It detects and 'latches onto' regions in the image that are generated by objects of 
interest in the environment. This results in deictic pointers or `anchors' that establish and monitor indexical references 
between internal symbols and external objects. Tracking not only takes place for a single object, but for 
an open-ended set of objects at different hierarchical levels, as long as they are part of the same spatio-temporal context.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth]{chap10/figs/peract-images}}
\caption{\label{fig:peract-images} 
Top: Some of the images recorded during the action of picking up a cylinder. 
Middle: Results of segmentation algorithms. 
Bottom: The spatio-temporal continuity of objects is tracked. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.70\textwidth]{chap10/figs/comparison}}
\caption{\label{fig:comparison} 
Comparison between two different strategies: The top-graph shows the growth in the average number of 
markers for the event-specific marker strategy. 
The bottom two graphs show the average size of the set of markers when analogy is used. As usual, the y-axis shows 
the number of games played. Overall, fewer markers are needed, making the case marking system more efficient 
and thus explaining why we get abstract semantic roles.}
\end{figure}

>>>>

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.50\textwidth]{chap10/figs/look-ball.pdf}
\includegraphics[width=.45\textwidth]{chap10/figs/views-ball.pdf}}
\caption{\label{f:look-ball} 
Left: Fr\'{e}d\'{e}ric Kaplan interacts with the AIBO robot in an experiment for the social learning of lexicons.
Right: View of the ball through the eye of the robot. As the same object is now seen from many different viewpoints it 
becomes much more challenging to consistently recognise it.}
\end{figure}
The first AIBO experiments did not involve a population of agents but rather a single agent that was interacting directly with a human
experimenter. The agent had to learn the lexicon of the human through social 
interaction within the context of situated language games. The 2001
paper ``AIBO's first words'' by Steels and Kaplan\todo{citation?} described experiments in which an AIBO robot 
learned words about objects in the environment, such as `ball' or `smiley', as well as words for actions, such as 
`sit down' or `stand up'.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.70\textwidth]{chap10/figs/fsm.pdf}}
\caption{\label{fig:fsm} 
The behaviour on the AIBO, including the language game scripts, are programmed using probabilistic finite state 
machines that bring the robot from one internal state to another one depending on internal or external conditions.}
\end{figure}

@@

The Perspective Reversal Experiment applied this methodology to explain a remarkable universal feature 
of human languages, namely, 
that speakers may use a different perspective on the scene than their own when 
conceptualising what to say, and that they may explicitly mark perspective switching, using lexical or grammatical means. Suppose
a speaker is facing a hearer, then it indeed makes a big difference whether she says 
\emph{the door to your left}  or \emph{the door to my left}. If the speaker says
\emph{the door to my left} she expects the hearer to perform an egocentric
perspective transformation and see the situation from her own point of view. Perspective switching is here signalled by 
explicitly expressing which perspective is used: {\itshape your} left versus {\itshape my} left. 
The German prepositions \emph{herein} and \emph{hinein}, both meaning `into' are another example of perspective marking. They
distinguish whether the direction of movement is 
towards the speaker \emph{herein} or away from the speaker \emph{hinein}, analogous to English \emph{come} (to where I am) 
versus \emph{go} (to another location). Although there are obvious differences in how languages express
perspective, there can be no doubt that perspective marking is pervasive across 
languages and language families, and that speakers make abundant use of it.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap10/figs/tokyo-persp.pdf}}
\caption{\label{fig:tokyo-persp} 
Luc Steels demonstrates live the Perspective Reversal Experiment in Tokyo, April 2005. AIBO robots track 
the movement of a ball and 
the robot acting as speaker describes to the hearer an observed movement, possibly taking into account the 
perspective of the hearer. For example, the speaker might say `the ball rolls from my left to your right'.}
\end{figure}

The Perspective Reversal Experiment worked as follows. A population of AIBO robots roam around
freely in an unconstrained in-door environment. As soon as one sees the
ball, it comes to a stop and searches for another robot nearby, which also
looks for the ball and stops when it sees it. Then the human
experimenter pushes the ball with a stick so that it rolls a short
distance (see \figref{fig:tokyo-persp}). The two robots next play a description game. One robot (acting as the `speaker') 
describes the ball-moving event to the other robot (the ``hearer''), going through the same semiotic cycle as the Talking 
Heads agents. The game is a success if the meaning conveyed by the speaker is compatible with the hearer's 
own perception of the scene.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.90\textwidth]{chap10/figs/persp-rev-perception.pdf}}
\caption{\label{fig:persp-rev-perception} 
Chain from visual image (left) to an analog world model (middle) and a discrete event description
that is the topic of the language game (right). 
The top row shows this chain for robot A and the bottom row the chain for robot B, both looking at the same ball.}
\end{figure}

Within this experimental setting three different strategies were tried in a long-term experiment where
a new strategy was injected at certain stages and performance monitored: 
\begin{itemize}
\item In stage I, agents use 
the Discrimination Game strategy and the Naming Game strategy discussed in chapters 4 and 5. A 
lexicon is emerging and there are some successful games (those for which the speaker and the hearer see the situation from 
the same perspective). But the lexicon keeps expanding and success is very limited as games where robots have a different 
perspective on the rolling ball all fail. 
\item In stage II, agents have an extended 
strategy where agents can geometrically transform a model of the scene, based on their own perspective, into a model of the scene 
as seen from the perspective of the hearer. They can then conceptualise the scene from that perspective. 
Communicative success significantly increases and the 
lexicon starts to decrease. However there is ambiguity because the hearer cannot know whether
the speaker has conceptualised the 
scene from his own viewpoint or performed a perspective reversal, leading to a large number of failed communications 
(more than 50 \%). 
\item In stage III, a more elaborate stratergy is used. The speaker now expresses whether he has performed perspective 
reversal (as in `{\itshape to your} left' vs. simply `left'). Only with this strategy are agents able to reach high 
communicative success with a stable inventory. 
\end{itemize}

\begin{figure}[t]
\centerline{
\includegraphics[width=0.75\textwidth]{chap10/figs/stages.pdf}}
\caption{\label{fig:stages} 
Three strategies are compared. In Stage I (left) agents categorise the event from their 
own perspective. In Stage II (middle) they conceptualise the scene from the viewpoint of the other agent as well. In Stage III (right) 
they express that they performed a perspective reversal. Only the third strategy leads to successful performance.} 
\end{figure}

@@

The earliest attempts to investigate the emergence of syntax date back to 1997, in other words even before 
the first Talking Heads experiment started. The first report 
appeared as a paper in the Artificial Intelligence journal in 1998 \citep{Steels:1998}. 
It was described by the journal editors (Dan Bobrow and 
Mike Brady) as ``obviously in an early stage, but very provocative and suggestive.'' The paper was based on my keynote 
lecture at IJCAI 1997 in Nagoya and used a precursor of the pan-tilt cameras built by Tony Belpaeme. 
(see \figref{f:plate7}).

@@

``The grammar is seen as a natural continuation of the lexicon, in the sense that it consists also of associations between forms and meanings. Use and success are monitored for each association so that the same type of self-organised coherence arises in the group, as seen in the lexicon. The form is now a more complex structure, defined as a syntactic schema. The meaning is a semantic schema. The schemas
circumscribe a feature-set in terms of a set of slots, restrictions	
on the fillers of each slot, and constraints	on the combination	
of the fillers to form the total covered by the schema. \todo{Where does this quote end?}

@@

The categories restricting slot-fillers are either themselves defined in terms of schemas
(for example, syn-cat-77 could be the restriction on a slot-filler in another schema), or they
are defined as rules that are applied in a forward-chaining fashion during matching. Two
examples of rules related to the above schema are:
\begin{verbatim}
  rule 101: ([WORD (W U)]) => ([MEMBER syn-cat-761])
  rule 99: ([WORD (W O)]) => ([MEMBER Syn-Cat-751])
\end{verbatim}
Semantic schemas describe the language-specific semantic structures underlying the
meanings of complete word groups. The closest linguistic correspondent to a semantic
schema is the notion of a case-frame. The constraints indicate how the total meaning
is constructed/decomposed into the meaning of the parts. During interpretation such
constraints therefore perform the same role as Montague style semantic interpretation
functions. The slots correspond to cases such as agent, patient, time, distance, or arguments
of semantic functions. The categories used to constrain what can fill a slot correspond
in linguistic terms to selection restrictions like animate, human, edible, future, etc. The
schema has also an associated category for the whole so that hierarchical combination is
possible. An example of a semantic schema is:
\begin{verbatim}
Schema-542
  SLOTS: (sem-slot-51 sem-slot-50)
  DESCRIPTION-SET: 
       ([sem-slot-50 sem-cat-751] [sem-slot-51 sem-cat-761])
  CONSTRAINTS: 
       ((CONJUNCTION (>> sem-slot-50) (>> sem-slot-51)))
  CATEGORY: sem-cat-77
  USE: 10
  SUCCESS: 3
\end{verbatim}
Inference rules such as the following define the selection restrictions:
\begin{verbatim}
  rule 102: ([VISIB v-4111) => ([MEMBER sem-cat-761)
  rule 100: ([VER v-4311) => ([MEMBER sem-cat-751)
\end{verbatim}
Each association in the grammar couples a syntactic schema with a semantic schema.
The association can be used in two directions. If a syntactic schema is recognised, i.e., all its 
constraints are satisfied by the input utterance, the semantic schema is
used to reconstruct its meaning. If a semantic schema is recognised, the syntactic schema
is used to reconstruct the form. The association contains a mapping of the slots in order to
enable this reconstruction. For example, the association combining the above two schemas is: 
\begin{verbatim}
Association-271
  MEANING: Schema-542
  FORM : Schema-541
  MAPPING : ((syn-slot-51 sem-slot-51)
             (syn-slot-50 sem-slot-50))
  USE: 10
  SUCCESS: 3"
\end{verbatim}\citep{Steels:1998}

\begin{figure}[t]
\centerline{
\includegraphics[width=0.75\textwidth]{chap10/figs/associations.pdf}}
\caption{\label{fig:associations} 
Top: Two lexical associations have become active (association-259 and association-234) to cover the distinctive 
feature-set ([VISIB v-411] [VER v-431]) resulting from discrimination and perception. Bottom: 
The group of words and the meanings match with schemas forming part of a grammatical association that combines the words. }
\end{figure}
An application of these schemas in producing the utterance `(W O) (W U)' is represented in \figref{fig:associations} and shown 
here: 
\begin{verbatim}
127 ++> Speaker: head-40 Topic: 2 Context: 6 5 4 3 1 0
Categorial Perception:
  ([VISIB v-4111] [VER v-431])
  ([VER v-4311] [AREA v-4061])
Lexicon lookup: (Association-259 Association-234)
Syntactic structure:
  (syn-cat-77
    (syn-slot-50 (syn-cat-75 |(W O)|))
    (syn-slot-51 (syr-cat-76 |(W U)|)))
Semantic structure:
  (sem-cat-78
    (sem-slot-50 (sem-cat-75 (VER v-431)))
    (sem-slot-51 (sem-cat-76 (VISIB v-411))))
Meaning:
  ([VISIB v-4111] [VER v-4311)
Expression: ((W 0) (W U))
\end{verbatim}

@@

The paper also briefly described the mechanisms by which such a grammar emerged in the experiment in terms of a 
cognitive memory system:

\begin{quotation}
``The cognitive memory system acts in a first instant purely as a device that records 
a particular way in which language has been produced so that it can later be re-produced in the same way.
Also the hearer can perform this recording operation. He is presented with a specific set of word forms from which he can abstract a syntactic schema. He derives meaning from the definition of the words in the lexicon and the distinctive feature-set coming out of perception. From this, a semantic schema can be extracted.
More interesting grammars emerge when additional operations to restructure the set of grammatical associations are used. For example, in the case of a partial match, substructures may be recategorised so that they nevertheless fit, two partially overlapping schemes with common fillers may lead to a new schema that integrates both, etc.''
\end{quotation}

@@

Based on these components, I implemented an initial experiment with only two agents.\footnote{These experiments were first reported at the Harvard Evolution of Language conference in 2002 with a paper by Luc Steels
entitled ``Computer Simulations of the origins of Case grammar''. The paper was later deemed 
``too incomprehensible for the Evolution of Language community'' to be acceptable for the conference 
proceedings edited by Maggie Tallerman.}
In 2005 Remi van Trijp picked up this line of research again and worked out the experiment 
further with great skill and sophistication. He extended the 
2-agent simulation to a multi-agent simulation and could finally report a complete running system in his 2008 Ph.D 
thesis \citep{VanTrijp:2014}. Here I give only a brief account of the first attempts towards case grammar
and how this pushed the state of the art in language game research.

But first we need a clearer idea on what case grammar is.
In the sentence `John gave Mary a book', John plays the role of agent, the giver 
of the book, Mary that of recipient or 
beneficiary and the book is the ``patient'' or object of the action. Languages differ in terms of the strategies that they 
use to express these participant roles. Some of them, like German, have an elaborate system of morphological 
markers attached to constituents of the noun phrase. These markers signal that the 
phrase is nominative, accusative, dative, etc., and these cases then map in complex ways 
onto semantic roles such as agent and patient.
Other languages, like English, use a strategy based on word order and prepositions. For example, 
`Mary gave John a book' contains the same participants as `John gave Mary a book' but now Mary is the 
agent and John the recipient. Still other languages, like Japanese, use a strategy based on particles following 
the noun, as in the Japanese sentence `Tanaka-san wa Tokyo de o-to-san ni atta' 
meaning Tanaka met his father in Tokyo. Tanaka-san is the name of a person, oto-san means `father' and atta 
means `met'. The markers are `wa' for the topic, `de' for location and `ni' for another sense of location which can 
also function as possessive. The case grammar experiments focused on modeling the latter strategy.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.60\textwidth]{chap10/figs/peract-setup.pdf}}
\caption{\label{fig:peract-setup} 
The installation for the case grammar experiments used a more sophisticated vision system and objects that 
could be moved around. We see the two Talking Heads cameras in 
the background. They are oriented towards objects on the table or situations in the 
environment. The monitors show the camera images and the states of image processing, just as for the original 
Talking Heads experiment. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.40\textwidth]{chap10/figs/adam-eve.pdf}}
\centerline{\includegraphics[width=.70\textwidth]{chap10/figs/jack-give.pdf}}
\centerline{\includegraphics[width=.80\textwidth]{chap10/figs/push-jill.pdf}}
\caption{\label{fig:adam-eve} 
Top: Computer animation of the speaking and hearing agent used in public demonstrations 
of the case experiment. 
Middle: Example of a puppet scene, where Jill gives a block to Jack. It shows the source image (left) and a (partial) 
result of image processing (right). Bottom: Another example of a puppet scene where Jill pushes the block.}
\end{figure}

{\bfshape  Scaling up the vision system} 
The vision system of the Talking Heads only dealt with static scenes. This had to be scaled up to
deal with dynamical scenes and event recognition. It was the focal work of 
Jean-Christophe Baillie at the Sony Computer Science Laboratory in Paris, who built a vision system 
called PERACT \is{PERACT vision system} inspired by an event-recognition developed by Jeffrey 
Siskind.\footnote{The vision system is described in the Ph.D thesis of J-C Baillie. A summary is found here: 
\cite{Steels:2003}. Inspiration for the event recognition system came 
primarily from the paper by \cite{Siskind:2000}.}
The vision system was decomposed into three subsystems with information flowing both in a bottom-up
(from image to world model) and top-down way (from world model and predictions to image): the subsystems achieved 
object identification, event identification, and derivation of qualitative descriptors.

{\bfshape  Object Identification} The first subsystem attempts to detect and track visual units at different hierarchical levels. 
It detects and 'latches onto' regions in the image that are generated by objects of 
interest in the environment. This results in deictic pointers or ``anchors'' that establish and monitor indexical references 
between internal symbols and external objects. Tracking not only takes place for a single object, but for 
an open-ended set of objects at different hierarchical levels, as long as they are part of the same spatio-temporal context.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth]{chap10/figs/peract-images.pdf}}
\caption{\label{fig:peract-images} 
Top: Some of the images recorded during the action of picking up a cylinder. 
Middle: Results of segmentation algorithms. 
Bottom: The spatio-temporal continuity of objects is tracked. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.70\textwidth]{chap10/figs/comparison.pdf}}
\caption{\label{fig:comparison} 
Comparison between two different strategies: The top-graph shows the growth in the average number of 
markers for the event-specific marker strategy. 
The bottom two graphs show the average size of the set of markers when analogy is used. As usual, the y-axis shows 
the number of games played. Overall, fewer markers are needed, making the case marking system more efficient 
and thus explaining why we get abstract semantic roles.}
\end{figure}
